---
title: Import
output: html_notebook
---

Data Source: https://catalog.data.gov/dataset/nutrition-physical-activity-and-obesity-behavioral-risk-factor-surveillance-system

```{r setup}
pacman::p_load(janitor, knitr, rio, tidyr, tidyverse)
brfss <- import("archive.zip") |> clean_names()
cat(c("BRFSS Dimensions:", dim(brfss)))

census_regions_file <- "https://github.com/cphalpert/census-regions/raw/master/us%20census%20bureau%20regions%20and%20divisions.csv"
census_regions <- import(census_regions_file) |> clean_names()
```

# About

I used this RMarkdown document to explore the CDC BRFSS data set and manipulate it until I had a useful data set for class. I thought about extracting the code I needed to just reorganize the data into a separate script, but running this doesn't actually take that long so the effort seemed pointless.

I also import the US Census Regions and add these to the CDC data to help my students further explore the data. I used the Census regions as stored cphalpert at [census-regions](https://github.com/cphalpert/census-regions). I use this data set often, because, as he notes, PDFs suck.

# Column Names

```{r col_names}
names(brfss) |> knitr::kable()
```



# Year

```{r}
brfss |>
  select(year_start, year_end) |>
  distinct()
```

Since all the years begin/end in the same year, I will have a single year column called `year`.

```{r}
brfss <- brfss |>
  select(-year_end) |>
  rename(year = year_start)
cat(c("BRFSS Dimensions:", dim(brfss)))
```



# Location Abbr/Desc/ID

```{r}
brfss |>
  select(geo_location, location_abbr, location_desc, location_id) |>
  distinct()
```

- I don't need geo_location. Otherwise, these _are_ useful.
- I need to document it so that I remember to separate brfss and state upon import.

```{r}
brfss <- brfss |> select(-geo_location)
cat(c("BRFSS Dimensions:", dim(brfss)))
```



# Data Source

```{r}
brfss |>
  select(datasource) |>
  distinct()
```

And since it is all one, We don't need this. We know this comes from the BRFSS data.

```{r}
brfss <- brfss |> select(-datasource)
cat(c("BRFSS Dimensions:", dim(brfss)))
```



# Class (ID), Topic (ID), Question (ID)

```{r}
brfss |>
  select(class, class_id) |>
  distinct()
```

So this is useful, although it should be all together.



# Data Value Unit, Data Value Type (ID), Data Value, Data Value Alt, 

```{r}
brfss |>
  select(
    data_value_unit,
    data_value_type,
    data_value_type_id,
    data_value,
    data_value_alt
  ) |>
  distinct()
```

- Data Value clearly has some value.
-  I don't think we need Data Value Alt.
-  Data Value Unit is just NA, DROP.
-  Data Value Type and Data Value Type ID are useless.

```{r}
brfss <-
  brfss |>
  select(
    -data_value_unit,
    -data_value_type,
    -data_value_type_id,
    -data_value_alt
  )
cat(c("BRFSS Dimensions:", dim(brfss)))
```



# Data Value Footnote Symbol, Data Value Footnote

```{r}
brfss |>
  select(
    data_value_footnote,
    data_value_footnote_symbol
  ) |>
  distinct()
```

This is entirely useless.

```{r}
brfss <-
  brfss |>
  select(
    -data_value_footnote,
    -data_value_footnote_symbol
  )
cat(c("BRFSS Dimensions:", dim(brfss)))
```

I'm down to 25 columns!



# Low/High Confidence Limit

For the analysis I want to do, I'm simply going to use the estimated average.

```{r}
brfss <-
  brfss |>
  select(
    -low_confidence_limit,
    -high_confidence_limit
  )
cat(c("BRFSS Dimensions:", dim(brfss)))
```



# Sample Size

I know I want to keep this since it is actually valuable.



# Total

```{r}
brfss |>
  group_by(total) |>
  count()
```

- Sigh.
- This seems to actually be connected to the next section.



# Stratification Category 1, Stratification 1, Stratification Category 1, Stratification Category ID, Stratification ID

```{r}
brfss |>
  select(
    stratification_category1,
    stratification_category_id1,
    stratification1,
    stratification_id1,
    total
  ) |>
  distinct()
```

- Dear god. These are actually (mostly) useful.
- Total is not.
- But age_years, education, gender, income, & race_ethnicity are all not useful.
- Let me confirm that.

```{r}
brfss |>
  select(stratification1, age_years) |>
  distinct()

brfss |>
  select(stratification1, education) |>
  distinct()
```

Drop all of these columns.

```{r}
brfss <-
  brfss |>
  select(
    -age_years,
    -education,
    -gender,
    -income,
    -race_ethnicity,
    -total
  )
cat(c("BRFSS Dimensions:", dim(brfss)))
```

- Six more columns gone like the wind. Wow
- And I want to do some simple renaming.

```{r}
brfss <-
  brfss |>
  rename(
    stratification_category = stratification_category1,
    stratification_category_id = stratification_category_id1,
    stratification = stratification1,
    stratification_id = stratification_id1
  )
```

This doesn't remove any columns, but it makes the data easier to manage.

# Topic, Question, Question ID

It would be nice to reduce the number of rows dramatically.

```{r}
brfss |>
  select(topic_id, topic, question_id, question) |>
  distinct() |>
  kable()
```

- What I need to do is to create a separate data set with the questions data.
- And then create convenient column labels.
- And then remove everything except the question_id.

```{r}
questions <-
  brfss |>
  select(class_id, class, topic_id, topic, question_id, question) |>
  distinct()
brfss <- brfss |> select(-class, -class_id, -topic_id, -topic, -question)
cat(c("BRFSS Dimensions:", dim(brfss)))
```

And now . . . .

|topic_id |topic                            |question_id |question                                                                                                                                                                                                                                              | question_label                    |
|:--------|:--------------------------------|:-----------|:-----------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|:----------------------------------|
|OWS1     |Obesity / Weight Status          |Q036        |Percent of adults aged 18 years and older who have obesity                                                                                                                                                                                            | obese                             |
|PA1      |Physical Activity - Behavior     |Q047        |Percent of adults who engage in no leisure-time physical activity                                                                                                                                                                                     | no_physical_activity              |
|OWS1     |Obesity / Weight Status          |Q037        |Percent of adults aged 18 years and older who have an overweight classification                                                                                                                                                                       | overweight                        |
|FV1      |Fruits and Vegetables - Behavior |Q018        |Percent of adults who report consuming fruit less than one time daily                                                                                                                                                                                 | no_fruits                         |
|FV1      |Fruits and Vegetables - Behavior |Q019        |Percent of adults who report consuming vegetables less than one time daily                                                                                                                                                                            | no_vegetables                     |
|PA1      |Physical Activity - Behavior     |Q043        |Percent of adults who achieve at least 150 minutes a week of moderate-intensity aerobic physical activity or 75 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)                                                  | exercise_150                      |
|PA1      |Physical Activity - Behavior     |Q046        |Percent of adults who engage in muscle-strengthening activities on 2 or more days a week                                                                                                                                                              | muscle_strengthening              |
|PA1      |Physical Activity - Behavior     |Q044        |Percent of adults who achieve at least 150 minutes a week of moderate-intensity aerobic physical activity or 75 minutes a week of vigorous-intensity aerobic physical activity and engage in muscle-strengthening activities on 2 or more days a week | exercise_150_muscle_strengthening |
|PA1      |Physical Activity - Behavior     |Q045        |Percent of adults who achieve at least 300 minutes a week of moderate-intensity aerobic physical activity or 150 minutes a week of vigorous-intensity aerobic activity (or an equivalent combination)                                                 | exercise_300                      |

Which really boils down to . . . . 

|question_id | question_label                    |
|:-----------|:----------------------------------|
|Q036        | obese                             |
|Q047        | no_physical_activity              |
|Q037        | overweight                        |
|Q018        | no_fruits                         |
|Q019        | no_vegetables                     |
|Q043        | exercise_150                      |
|Q046        | muscle_strengthening              |
|Q044        | exercise_150_muscle_strengthening |
|Q045        | exercise_300                      |

```{r}
brfss <-
  brfss |>
  mutate(
    question_label = case_when(
      question_id == "Q036" ~ "obese",
      question_id == "Q047" ~ "no_physical_activity",
      question_id == "Q037" ~ "overweight",
      question_id == "Q018" ~ "no_fruits",
      question_id == "Q019" ~ "no_vegetables",
      question_id == "Q043" ~ "exercise_150",
      question_id == "Q046" ~ "muscle_strengthening",
      question_id == "Q044" ~ "exercise_150_muscle_strengthening",
      question_id == "Q045" ~ "exercise_300"
    )
  ) |>
  select(-question_id)

brfss_obese <-
  brfss |>
  filter(question_label == "obese") |>
  pivot_wider(names_from = question_label, values_from = data_value)
brfss_no_physical_activity <-
  brfss |>
  select(
    -location_abbr,
    -location_desc,
    -stratification_category,
    -stratification,
    -sample_size
  ) |>
  filter(question_label == "no_physical_activity") |>
  pivot_wider(names_from = question_label, values_from = data_value)
brfss_overweight <-
  brfss |>
  select(
    -location_abbr,
    -location_desc,
    -stratification_category,
    -stratification,
    -sample_size
  ) |>
  filter(question_label == "overweight") |>
  pivot_wider(names_from = question_label, values_from = data_value)
brfss_no_fruits <-
  brfss |>
  select(
    -location_abbr,
    -location_desc,
    -stratification_category,
    -stratification,
    -sample_size
  ) |>
  filter(question_label == "no_fruits") |>
  pivot_wider(names_from = question_label, values_from = data_value)
brfss_no_vegetables <-
  brfss |>
  select(
    -location_abbr,
    -location_desc,
    -stratification_category,
    -stratification,
    -sample_size
  ) |>
  filter(question_label == "no_vegetables") |>
  pivot_wider(names_from = question_label, values_from = data_value)
brfss_exercise_150 <-
  brfss |>
  select(
    -location_abbr,
    -location_desc,
    -stratification_category,
    -stratification,
    -sample_size
  ) |>
  filter(question_label == "exercise_150") |>
  pivot_wider(names_from = question_label, values_from = data_value)
brfss_muscle_strengthening <-
  brfss |>
  select(
    -location_abbr,
    -location_desc,
    -stratification_category,
    -stratification,
    -sample_size
  ) |>
  filter(question_label == "muscle_strengthening") |>
  pivot_wider(names_from = question_label, values_from = data_value)
brfss_exercise_150_muscle_strengthening <-
  brfss |>
  select(
    -location_abbr,
    -location_desc,
    -stratification_category,
    -stratification,
    -sample_size
  ) |>
  filter(question_label == "exercise_150_muscle_strengthening") |>
  pivot_wider(names_from = question_label, values_from = data_value)
brfss_exercise_300 <-
  brfss |>
  select(
    -location_abbr,
    -location_desc,
    -stratification_category,
    -stratification,
    -sample_size
  ) |>
  filter(question_label == "exercise_300") |>
  pivot_wider(names_from = question_label, values_from = data_value)
```

And now we try to get it all back together again . . .

```{r}
join_cols <-
  c("year", "location_id", "stratification_category_id", "stratification_id")

brfss <-
  brfss_obese |>
  left_join(
    census_regions,
    by = c("location_abbr" = "state_code", "location_desc" = "state")
  ) |>
  left_join(brfss_overweight, by = join_cols) |>
  left_join(brfss_no_physical_activity) |>
  left_join(brfss_exercise_150) |>
  left_join(brfss_exercise_300) |>
  left_join(brfss_muscle_strengthening) |>
  left_join(brfss_exercise_150_muscle_strengthening) |>
  left_join(brfss_no_fruits) |>
  left_join(brfss_no_vegetables) |>
  select(
    year,
    location_id,
    location_abbr,
    location_desc,
    region,
    division,
    stratification_category_id,
    stratification_category,
    stratification_id,
    stratification,
    sample_size,
    obese,
    overweight,
    no_physical_activity,
    exercise_150,
    exercise_300,
    muscle_strengthening,
    exercise_150_muscle_strengthening,
    no_fruits,
    no_vegetables
  ) |>
  arrange(year, location_id, stratification_category_id, stratification_id)
```

And now I just need to save it all to disk.

```{r}
write_csv(questions, "questions.csv", na = "")
write_csv(brfss, "brfss.csv", na = "")
```